Map{R}educe in {MPI} for Large-scale Graph Algorithms

### 2010-11-03 paper/Response.tex

\setlength{\voffset}{-0.0in}
\hoffset=-50pt
\topmargin=0pt
\headheight=0pt
\headsep=0pt
\evensidemargin=72pt % 1in+7pt

\setlength{\parskip}{0.5ex}
\setlength{\parindent}{1em}
\setlength{\floatsep}{1pt}
\setlength{\textfloatsep}{2pt}
\setlength{\intextsep}{2pt}

\begin{document}

October 29, 2010\\

Response to reviewers:  \\

We thank the reviewers for their careful reading of our paper.  Their
suggestions definitely strengthen the paper and sparked new ideas for
improving our algorithms.  Below we list the Reviewers' comments
verbatim.  Our responses are shown in italics. \\

Steve Plimpton and Karen Devine

@@

\vspace{0.5 in}

Reviewer \#1: The submitted paper features a description and analysis of a
new MapReduce implementation based on MPI.  The features and limitations
of the implementation are covered with respect to other MapReduce
implementations.  The implementation is applied to several
graph-theoretic algorithms, and the performance of these applications
is compared with applications from an established library (Trilinos).

The introductory and supporting text are, overall, very good.  The
paper places MapReduce in its historical context and motivates its
usage.  The outline of strengths is helpful- important features
include support for out-of-core operations that other libraries (e.g.,
Trilinos) often cannot handle and the ability to use MPI directly in
addition to the MapReduce abstraction.  The paper then provides a good
overview of the composition of operations in a MapReduce application
and a summary of the out-of-core mechanisms available.  The figure
(Figure 1) used to provide a high-level schematic of MapReduce data
flows could be enhanced- for example, it should indicate node-local
and communication intensive operations.

{\it Figure 1 is actually describing an on-processor rearrangement of
data that each processor performs on its local data it owns.  Thus
there is no inter-processor communication involved.}

@@

All performance results are shown as scaling studies for a range of
problem sizes for the six graph algorithms.  The results show that the
MapReduce applications do not compete well with the reference
implementation (Trilinos), but the authors make clear the limitations
involved in using this high-level library which trades some
performance for usability.  Additionally, the implementation supports
very large data sets which can only be processed using its out-of-core
feature set.  The results are presented well.

Since improved performance for small and mid-range problems is not the
contribution of the paper, it could be improved by stressing other
aspects of the system.  Real-world out-of-core problems could be
surveyed.  Code size and complexity could be considered with usability
as a motivation.  Is the MapReduce-based graph algorithm library
easier to use than the Trilinos library?  Should higher-level user
applications be considered?  Can the MapReduce-based inefficiencies be
isolated more concisely?

The paper does not have a concise "Conclusion", only a "Lessons Learned"
section.

{\it We have made some of these comparison points more
clearly in the final ``Lessons Learned'' section.}

%comment on MR vs Trilinos, even if libraries already exist
%  cannot use linear algebra lib for all graph algorithms described
%  here - PR is an exception since can be cast as matvec
%added line counts of algorithms

@@

\vspace{0.5 in}

Reviewer \#2: This is a nice paper addressing an important current topic
with a good quality open-source implementation.  The overall presentation
is good: the description of algorithms and implementations is clear and the
performance evaluation is detailed. As the authors pointed out, the
performance of the MR-MPI implementations of some graph algorithms falls
behind the matrix-based counterparts so it would be interesting to further
investigate where the performance descrepency comes from and if it's
possible to optimize the algorithm and implementation to bridge the
performance gap.

{\it We instrumented our MR-MPI and Trilinos implementations of
PageRank to show where time is spent during the PageRank iterations.
These results were added in Table 2.  We further instrumented the two
most expensive steps of the PageRank iterations, as shown in Tables 3
and 4.  These experiments showed that considerable time is spent in
reorganizing data from key-value pairs to key-multivalue pairs.  In
sections 4.1 and 5.2, we describe alternative data layouts that could
reduce the number of times the data must be reorganized to improve
performance.}

@@

\vspace{0.5 in}

Reviewer \#3: I appreciated the clear exposition of your map/reduce
framework, its decomposition into a number of independent and composable
stages, and the explanation of how out-of-core computations are handled.

I also liked that you presented the map/reduce implementation of a
number of different graph algorithms. However, I thought the
presentation of these algorithms could be improved in two ways: - The
text tends to just explaining what the code does, not why, nor why
this solves the specific problem (4.2 was the worst culprit here - f I
understood correctly, the vertices get labeled by edge counts that are
used to give the edges a direction that helps identify triangles: if
this is correct a little diagram showing this would help...)  - The
detailed algorithms (Figures 4-9) are a little too imprecise to follow
at times. In particular, when multiple map/reduces are involved the
lack of names for the map/reduce operations and their inputs makes
things confusing. Please be very explicit (after all the code itself
is), and Figure 4 at least actually had names for the map-reduce
objects.

{\it We significantly re-worked and standardized the notation used
in Section 4 to describe all the graph algorithms.  In particular
we noted which operations are performed with which MapReduce
objects and their associated key/value data, so that the flow
of data through each algorithm should be more clear.  A discussion
of the MR-MPI library's ability to allow creation of multiple
MapReduce objects was added to Section 2.}

### 2010-11-03 paper/graph.tex
Minor change
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@646 3fc257bb-8550-0410-81b3-a5c103da83f7
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@655 3fc257bb-8550-0410-81b3-a5c103da83f7

We begin with a MapReduce procedure for creating large, sparse,
randomized graphs, since they are the input for the algorithms
discussed below.  R-MAT matrices ~\cite{RMAT} are recursively
generated graphs with power-law degree distributions.  They are
commonly used to represent web and social networks.  The user
specifices six parameters that define the graph: the number of
vertices $N$ and edges $M$, and four parameters $a$, $b$, $c$, $d$
that sum to 1.0 and are discussed below.  The algorithm in
Figure~\ref{fig:rmat} generates $M$ unique non-zero entries in a
sparse $N \times N$ matrix $A$, where each entry $A_{ij}$ represents
an edge between graph vertices $(V_i,V_j)$.

@@

 \caption{\it MapReduce algorithm for R-MAT graph generation on $P$
 processors.}

@@

 \caption{\it Enhanced MapReduce algorithm for R-MAT graph generation
 on $P$ processors.}

@@

The algorithm in Figure~\ref{fig:pr2} performs these iterations, using
linear-algebra concepts implemented in a MapReduce framework.  The
nonzeros of matrix $A$ are initialized as described above; the
PageRank vector $x$ is initialized uniformly.  A MapReduce object $MT$
contains the indices of all-zero rows of $A$, corresponding to
vertices with no outlinks.  The PageRank algorithm is made more
efficient by calling {\it aggregate()} once on the MapReduce objects
representing the matrix $A$, the all-zero matrix rows $MT$, and the
vector $x$ before the PageRank iterations begin.  Since these three
MapReduce objects have the same key types, pre-aggregating the objects
moves all KV pairs with a given key to a single processor.  Thus, many
of the PageRank operations become local operations, replacing {\it
collate()} operations (which are equivalent to an {\it aggregate()}
followed by {\it convert()}) with {\it convert()} calls.  In
Figure~\ref{fig:pr2}, steps where {\it convert()} operations can be
substituted for {\it collate()} operations are marked with an
asterisk.  This strategy of pre-aggregating to improve data locality
is useful in many MapReduce algorithms, but assumes that specific keys
are always mapped to the same processor.

In Step 1 of a PageRank interation, $MT$ is used in dot products to
compute adjustments to the PageRank vector to represent the uniform
probability of going from a leaf page to any other page.  Adjustments
for random jumps from any page are also computed.  In Step 2, the
matrix-vector product $A^T x$ is computed.  This is the most expensive
part of the computation, requiring two passes over the nonzero
structure of $A$.  In the first pass, a {\it convert()} gathers all
local row entries $A_{ij}$ with their associated $x_i$ entry, and a
{\it reduce()} computes $A_{ij} x_i$.  In the second pass, a {\it
collate()} gathers, for each $j$, all contributions to the column sum
$\sum_i A_{ij} x_i$, which is computed by a second {\it reduce()}.
Steps 3 and 4 adjust and scale the product vector; the work is
proportional to $|V|$.  Step 5 computes the residual to determine
whether the PageRank vector has converged; work is again proportional
to $|V|$.  In Step 6, the PageRank vector is overwritten prior to the
next iteration.  Throughout the algorithm, we exploit the ability to
call MPI\_Allreduce to compute global norms and residuals.

This linear-algebra approach is attractive because it allows construction
of the MapReduce PageRank algorithm from simple operations (matrix-vector
multiplication, dot products) on commonly used data types (matrices, vectors).
One can build a variety of algorithms based on these primitives;
indeed, linear-algebra toolkits such as Trilinos~\cite{Trilinos-Overview} and 
PETSc~\cite{petsc-efficient} use this exact design.  In parallel, these toolkits
share pre-computed data layouts and communication maps among distributed
objects to enable efficient operations between the objects.
For MapReduce, however, 
separating the data into vectors and matrices results in additional overhead 
to gather key-value pairs with the same key into a key-multivalue pair.  
This overhead can be reduced in PageRank, for example, by storing the
$x$ and $y$ vectors in a single MapReduce object, with key $i$ and value 
$(x_i, y_i)$.  Doing so eliminates the copy operation in Step 2 of 
Figure~\ref{fig:pr2}, as well as the add and convert operations in Step 5.
A further performance improvement would store an entire row of $A$ along with
$x$ and $y$ in a single MapReduce object, with key $i$ and value 
$(x_i, y_i, d_{out}, [j_1, A_{ij_1}], [j_2, A_{ij_2}], \dots, [j_{d_{out}}, 
A_ij_{d_{out}}])$ (that is, storing together, for each vertex, its old and
new PageRank value along with all out-going edges incident to the vertex).  
This data layout may further reduce time spent in converting from key-value 
pairs to key-multivalue pairs, but at the cost of code reuse and scalability.  
While this data layout may be efficient for PageRank, it is more complex
and less likely to be useful for additional algorithms.  In addition, since
each key-value pair must fit within a single page in MR-MPI, storing entire
matrix rows in a single key-value pair limits the size of the the matrices
on which the algorithm can be applied.  The more general approach of storing
non-zeros separately enables arbitrarily large problems to be solved.
The detailed timing break-downs in section~\ref{subsec:results_pagerank} 
allow one to estimate the benefit of these implementations relative to the
code complexity and scalability.

%\begin{figure}[htb]
% \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
%  xxxx\=xxxxx\=xxxxx\=xxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill
%Input: \\
%\> MapReduce object $A$ containing $|V| \times |V|$ matrix:  Key = $i$, Value = $[j, A_{ij}]$ \\
%\> MapReduce object $MT$ containing indices of all-zero rows of $A$:  Key = $i$, Value = 0 \\
%\> MapReduce object $x$ containing initial PageRank vector:  Key = $i$, Value = $1/|V|$ \\
%While (${residual} > {tolerance}$) \\
%\> 1. Compute contribution for random jumps and zero-outdegree vertices \\
%\> \> Add: $x$ to $MT$ \\
%\> \> Collate $MT$:  Vertex is Key \\
%\> \> Reduce $MT$:   \\
%\> \> \> Input:  Key = $i$; MultiValue = $x_i, 0$ if $i \in {MT}$; MultiValue = $x_i$ if $i \notin {MT}$ \\
%\> \> \> Compute:  Local contributions $c_{local}$ \\
%\> \> \> Output:  If ${nvalues} = 2$, Key = $i$; Value = 0 \\
%\> \> MPI\_Allreduce $c_{local}$ with MPI\_SUM to compute global contribution $c_{global}$ \\
%\> 2. Compute $y = A^T x$ \\
%\> \> Copy $x$ to $y$ \\
%\> \> Add $A$ to $y$ \\
%\> \> Collate $y$:  Key is row index of $A$ \\
%\> \> Reduce $y$: \\
%\> \> \>  Input: Key = $i$; Multivalue = $i, [j, A_{ij}]$ for nonzeros $A_{ij}$ in $A$ \\
%\> \> \>  Output:  Key = $j$; Value = $A_{ij} y_i$ \\
%\> \> Collate $y$:  Key is column index of $A$ \\
%\> \> Reduce $y$:  \\
%\> \> \> Input:  Key = $j$; Multivalue = $A_{ij} y_i$ for all non-zero entries $A_{ij}$ of column $j$ \\
%\> \> \> Output:  Key = $j$; Value = $\sum_i A_{ij} y_i$ \\
%\> 3. Add contribution $c_{global}$ to $y$; compute local max norm $n_{local}$ of $y$ \\
%\> \> Map $y$:  \\
%\> \> \> Input:  Key = $i$, Value = $y_i$ \\
%\> \> \> Compute: $y_i = y_i + c_{global}$; if $y_i > n_{local}$, $n_{local} = y_i$ \\
%\> \> \> Output:  Key = $i$, Value = $y_i$ \\
%\> \> MPI\_Allreduce $n_{local}$ with MPI\_MAX to compute global max norm $n_{global}$ \\
%\> 4. Scale $y$ by $n_{global}$ \\
%\> \> Map $y$: \\
%\> \> \> Input:  Key = $i$, Value = $y_i$ \\
%\> \> \> Output:  Key = $i$, Value = $y_i / {n_{global}}$ \\
%\> 5. Compute residual \\
%\> \> Add $y$ to $x$ \\
%\> \> Collate $x$:  Key is index $i$ \\
%\> \> Reduce $x$: \\
%\> \> \> Input:  Key = $i$; Multivalue = $x_i, y_i$ \\
%\> \> \> Compute:  if $|x_i - y_i| > {resid}$, ${resid} = |x_i - y_i|$ \\
%\> \> MPI\_Allreduce with MPI\_MAX to compute $residual$ \\
%\> 6. $x = y$; \\
%Output:  PageRank vector $x$
  %\end{tabbing}
 %\end{minipage}}\end{center}
%
 %\caption{\it MapReduce algorithm for PageRank vertex ranking.}
%
 %\label{fig:pr}
%\end{figure}

@@

Input: \\
$*$\> Pre-aggregated MapReduce object $A$ containing $|V| \times |V|$ matrix:  Key = $i$, Value = $[j, A_{ij}]$ \\
$*$\> Pre-aggregated MapReduce object $MT$ containing indices of all-zero rows of $A$:  Key = $i$, Value = 0 \\
$*$\> Pre-aggregated MapReduce object $x$ containing initial PageRank vector:  Key = $i$, Value = $1/|V|$ \\
While (${residual} > {tolerance}$) \\
\> 1 Compute contribution for random jumps and zero-outdegree vertices \\
\> \> Add: $x$ to $MT$ \\
$*$\> \> Convert $MT$:  Vertex is Key \\
\> \> Reduce $MT$:   \\
\> \> \> Input:  Key = $i$; MultiValue = $x_i, 0$ if $i \in {MT}$; MultiValue = $x_i$ if $i \notin {MT}$ \\
\> \> \> Compute:  Local contributions $c_{local}$ \\
\> \> \> Output:  If ${nvalues} = 2$, Key = $i$; Value = 0 \\
\> \> MPI\_Allreduce $c_{local}$ with MPI\_SUM to compute global contribution $c_{global}$ \\
\> 2 Compute $y = A^T x$  \\
\> \> Copy $x$ to $y$ \\
\> \> Add $A$ to $y$ \\
$*$\> \> Convert $y$:  Key is row index of $A$ \\
\> \> Reduce $y$: \\
\> \> \>  Input: Key = $i$; Multivalue = $y_i, [j, A_{ij}]$ for nonzeros $A_{ij}$ in $A$ \\
\> \> \>  Output:  Key = $j$; Value = $A_{ij} y_i$ \\
\> \> Collate $y$:  Key is column index of $A$ \\
\> \> Reduce $y$:  \\
\> \> \> Input:  Key = $j$; Multivalue = $A_{ij} y_i$ for all non-zero entries $A_{ij}$ of column $j$ \\
\> \> \> Output:  Key = $j$; Value = $\sum_i A_{ij} y_i$ \\
\> 3 Add contribution $c_{global}$ to $y$; compute local max norm $n_{local}$ of $y$ \\
\> \> Map $y$:  \\
\> \> \> Input:  Key = $i$, Value = $y_i$ \\
\> \> \> Compute: $y_i = y_i + c_{global}$; if $y_i > {n_{local}}$, $n_{local} = y_i$ \\
\> \> \> Output:  Key = $i$, Value = $y_i$ \\
\> \> MPI\_Allreduce $n_{local}$ with MPI\_MAX to compute global max norm $n_{global}$ \\
\> 4 Scale $y$ by $n_{global}$ \\
\> \> Map $y$: \\
\> \> \> Input:  Key = $i$, Value = $y_i$ \\
\> \> \> Output:  Key = $i$, Value = $y_i / {n_{global}}$ \\
\> 5 Compute residual \\
\> \> Add $y$ to $x$ \\
$*$\> \> Convert $x$:  Key is index $i$ \\
\> \> Reduce $x$:   \\
\> \> \> Input:  Key = $i$; Multivalue = $x_i, y_i$ \\
\> \> \> Compute:  if $|x_i - y_i| > {resid}$, ${resid} = |x_i - y_i|$ \\
\> \> MPI\_Allreduce with MPI\_MAX to compute $residual$ \\
\> 6 $x = y$; \\
Output: PageRank vector $x$
  \end{tabbing}
 \end{minipage}}\end{center}

 \caption{\it MapReduce algorithm for PageRank vertex ranking;
pre-aggregating the MapReduce objects allows many communication
intensive {\it collate()} operations to be replaced by strictly local
{\it convert()} operations (as marked with asterisks).}

@@

A triangle in a graph is any triplet of vertices $(V_i,V_j,V_k)$ where
the edges $(V_i,V_j), (V_j,V_k), (V_i,V_k)$ exist.  Figure
\ref{fig:tri} outlines a MapReduce algorithm that enumerates all
triangles, assuming an input graph of undirected edges $(V_i,V_j)$
where $V_i < V_j$ for every edge, i.e. an upper-triangular R-MAT
matrix.  This exposition follows the triangle-finding algorithm
presented in~\cite{Cohen09}.

@@

1 Copy: \> $G_0$ = copy of edge KV pairs from input graph \\
1 Map: \> Convert edges to vertices \\
           \> \> input Key = $(V_i,V_j)$, Value = NULL \\
           \> \> output Key = $V_i$, Value = $V_j$ \\
           \> \> output Key = $V_j$, Value = $V_i$ \\
1 Collate \\
1 Reduce: \> Add first degree to one vertex in edge \\
              \> \> input Key = $V_i$, MultiValue = $(V_j, V_k, ...)$ \\
	      \> \> for each V in MultiValue: \\
              \> \> \> if $V_i < V$: output Key = $(V_i,V)$, Value = $(D_i,0)$ \\
              \> \> \> else: output Key = $(V,V_i)$, Value = $(0,D_i)$ \\
2 Collate \\
2 Reduce: \> Add second degree to other vertex in edge \\
              \> \> input Key = $(V_i,V_j)$, MultiValue = $((D_i,0),(0,D_j))$ \\
              \> \> output Key = $(V_i,V_j)$, Value = $(D_i,D_j)$ with $V_i < V_j$ \\
3 Map: \> Low degree vertex emits edges \\
           \> \> if $D_i < D_j$: output Key = $V_i$, Value = $V_j$ \\
           \> \> else if $D_j < D_i$: output Key = $V_j$, Value = $V_i$ \\
           \> \> else: output Key = $V_i$, Value = $V_j$ \\
3 Collate \\
3 Reduce: \> Emit angles of each vertex \\
              \> \> input Key = $V_i$, MultiValue = $(V_j, V_k, ...)$ \\
	      \> \> for each $V_1$ in MultiValue: \\
	      \> \> \> for each $V_2$ beyond $V_1$ in MultiValue: \\
	      \> \> \> \> if $V_1 < V_2$: output Key = $(V_1,V_2)$, Value = $V_i$ \\
	      \> \> \> \> else: output Key = $(V_2,V_1)$, Value = $V_i$ \\
4 Add: \> Add $G_0$ edge KV pairs to angle KV pairs \\
4 Collate \\
4 Reduce: \> Emit triangles \\
              \> \> input Key = $(V_i,V_j)$, MultiValue = $(V_k,V_l,NULL,V_m,...)$ \\
              \> \> if NULL exists in MultiValue: \\
	      \> \> \> for each non-NULL V in MultiValue: \\
	      \> \> \> \> output Key = $(V_i,V_j,V)$, Value = NULL

>>>>

We begin with a MapReduce procedure for creating large, sparse,
randomized graphs, since they are the input for the algorithms
discussed below.  R-MAT matrices ~\cite{RMAT} are recursively
generated graphs with power-law degree distributions.  They are
commonly used to represent web and social networks.  The user
specifices six parameters that define the graph: the number of
vertices $N$ and edges $M$, and four parameters $a$, $b$, $c$, $d$
that sum to 1.0 and are discussed below.  The algorithm in
Figure~\ref{fig:rmat} uses a single MapReduce object to generate $M$
unique non-zero entries in a sparse $N \times N$ matrix $A$, where
each entry $A_{ij}$ represents an edge between graph vertices
$(V_i,V_j)$.

@@

 \caption{\it MapReduce algorithm for R-MAT graph generation on $P$
 processors, using a single MapReduce object.}

@@

 \caption{\it Enhanced MapReduce algorithm for R-MAT graph generation
 on $P$ processors, using two MapReduce objects $E_{new}$ and
 $E_{old}$.}

@@

The algorithm in Figure~\ref{fig:pr2} performs these iterations, using
linear-algebra concepts implemented in a MapReduce framework with data
stored in 4 MapReduce objects: $A$, $MT$, $x$, and $y$.  $A$ stores
the matrix, $MT$ its ``empty'' rows, while $x$ and $y$ are vectors.
The nonzeros of matrix $A$ are initialized as described above; the
PageRank vector $x$ is initialized uniformly.  A MapReduce object $MT$
contains the indices of all-zero rows of $A$, corresponding to
vertices with no outlinks.  The PageRank algorithm is made more
efficient by calling {\it aggregate()} once on the MapReduce objects
representing the matrix $A$, the all-zero matrix rows $MT$, and the
vector $x$ before the PageRank iterations begin.  Since these three
MapReduce objects have the same key types, pre-aggregating the objects
moves all KV pairs with a given key to a single processor.  Thus, many
of the PageRank operations become local operations, replacing {\it
collate()} operations (which are equivalent to an {\it aggregate()}
followed by {\it convert()}) with simply {\it convert()} calls.  In
Figure~\ref{fig:pr2}, steps where {\it convert()} operations were
substituted for {\it collate()} operations are marked with an
asterisk.  This strategy of pre-aggregating to improve data locality
is useful in many MapReduce algorithms, but assumes that specific keys
can always be mapped to the same processor.

In Step 1 of a PageRank interation, $MT$ is used in dot products to
compute adjustments to the PageRank vector to represent the uniform
probability of going from a leaf page to any other page.  Adjustments
for random jumps from any page are also computed.  In Step 2, the
matrix-vector product $A^T x$ is computed and the result storey in
$y$.  This is the most expensive part of the computation, requiring
two passes over the nonzero structure of $A$.  In the first pass, a
{\it convert()} gathers all local row entries $A_{ij}$ with their
associated $x_i$ entry, and a {\it reduce()} computes $A_{ij} x_i$.
In the second pass, a {\it collate()} gathers, for each $j$, all
contributions to the column sum $\sum_i A_{ij} x_i$, which is computed
by a second {\it reduce()}.  Steps 3 and 4 adjust and scale the
product vector; the work is proportional to $|V|$.  Step 5 computes
the residual to determine whether the PageRank vector has converged;
work is again proportional to $|V|$.  In Step 6, the PageRank vector
is overwritten prior to the next iteration.  Throughout the algorithm,
we exploit the ability to call MPI\_Allreduce to compute global norms
and residuals.

This linear-algebra approach is attractive because it allows
construction of the MapReduce PageRank algorithm from simple
operations (matrix-vector multiplication, dot products) on commonly
used data types (matrices, vectors).  One can build a variety of
algorithms based on these primitives; indeed, linear-algebra toolkits
such as Trilinos~\cite{Trilinos-Overview} and
PETSc~\cite{petsc-efficient} exploit this idea.  In parallel, these
toolkits share pre-computed data layouts and communication maps among
distributed objects to enable efficient operations between the
objects.  For MapReduce, however, separating the data into vectors and
matrices results in additional overhead to gather KV pairs with the
same key into a key-multivalue (KMV) pair.  This overhead can be
reduced in PageRank, for example, by storing the $x$ and $y$ vectors
in a single MapReduce object, with key $i$ and value $(x_i, y_i)$.
Doing so eliminates the {\it copy()} operation in Step 2 of
Figure~\ref{fig:pr2}, as well as the {\it add()} and {\it convert()}
operations in Step 5.  A further performance improvement would store
an entire row of $A$ along with $x$ and $y$ in a single MapReduce
object, with key $i$ and value $(x_i, y_i, d_{out}, [j_1, A_{ij_1}],
[j_2, A_{ij_2}], \dots, [j_{d_{out}}, A_ij_{d_{out}}])$ (that is,
storing together, for each vertex, its old and new PageRank value
along with all out-going edges incident to the vertex).  This data
layout eliminates an additional {it convert()} operation, but at the
cost of code reuse and scalability.  While this data layout may be
efficient for PageRank, it is more complex and less likely to be
useful for additional algorithms.  In addition, since each KV pair
must fit within a single page in MR-MPI, storing entire matrix rows in
a single KV pair limits the size of the the matrices on which the
algorithm can be applied.  The more general approach of storing
non-zeros separately enables arbitrarily large problems to be solved.
The detailed timing break-downs in
section~\ref{subsec:results_pagerank} allow one to estimate the
benefit of these implementations relative to code complexity and
scalability issues.

%\begin{figure}[htb]
% \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
%  xxxx\=xxxxx\=xxxxx\=xxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill

%\> MapReduce object $A$ containing $|V| \times |V|$ matrix:  Key = $i$, Value = $[j, A_{ij}]$ \\
%\> MapReduce object $MT$ containing indices of all-zero rows of $A$:  Key = $i$, Value = 0 \\
%\> MapReduce object $x$ containing initial PageRank vector:  Key = $i$, Value = $1/|V|$ \\
%While (${residual} > {tolerance}$) \\
%\> 1. Compute contribution for random jumps and zero-outdegree vertices \\
%\> \> Add: $x$ to $MT$ \\
%\> \> Collate $MT$:  Vertex is Key \\
%\> \> Reduce $MT$:   \\
%\> \> \> Input:  Key = $i$; MultiValue = $x_i, 0$ if $i \in {MT}$; MultiValue = $x_i$ if $i \notin {MT}$ \\
%\> \> \> Compute:  Local contributions $c_{local}$ \\
%\> \> \> Output:  If ${nvalues} = 2$, Key = $i$; Value = 0 \\
%\> \> MPI\_Allreduce $c_{local}$ with MPI\_SUM to compute global contribution $c_{global}$ \\
%\> 2. Compute $y = A^T x$ \\
%\> \> Copy $x$ to $y$ \\
%\> \> Add $A$ to $y$ \\
%\> \> Collate $y$:  Key is row index of $A$ \\
%\> \> Reduce $y$: \\
%\> \> \>  Input: Key = $i$; Multivalue = $i, [j, A_{ij}]$ for nonzeros $A_{ij}$ in $A$ \\
%\> \> \>  Output:  Key = $j$; Value = $A_{ij} y_i$ \\
%\> \> Collate $y$:  Key is column index of $A$ \\
%\> \> Reduce $y$:  \\
%\> \> \> Input:  Key = $j$; Multivalue = $A_{ij} y_i$ for all non-zero entries $A_{ij}$ of column $j$ \\
%\> \> \> Output:  Key = $j$; Value = $\sum_i A_{ij} y_i$ \\
%\> 3. Add contribution $c_{global}$ to $y$; compute local max norm $n_{local}$ of $y$ \\
%\> \> Map $y$:  \\
%\> \> \> Input:  Key = $i$, Value = $y_i$ \\
%\> \> \> Compute: $y_i = y_i + c_{global}$; if $y_i > n_{local}$, $n_{local} = y_i$ \\
%\> \> \> Output:  Key = $i$, Value = $y_i$ \\
%\> \> MPI\_Allreduce $n_{local}$ with MPI\_MAX to compute global max norm $n_{global}$ \\
%\> 4. Scale $y$ by $n_{global}$ \\
%\> \> Map $y$: \\
%\> \> \> Input:  Key = $i$, Value = $y_i$ \\
%\> \> \> Output:  Key = $i$, Value = $y_i / {n_{global}}$ \\
%\> 5. Compute residual \\
%\> \> Add $y$ to $x$ \\
%\> \> Collate $x$:  Key is index $i$ \\
%\> \> Reduce $x$: \\
%\> \> \> Input:  Key = $i$; Multivalue = $x_i, y_i$ \\
%\> \> \> Compute:  if $|x_i - y_i| > {resid}$, ${resid} = |x_i - y_i|$ \\
%\> \> MPI\_Allreduce with MPI\_MAX to compute $residual$ \\
%\> 6. $x = y$; \\
%Output:  PageRank vector $x$
  %\end{tabbing}
 %\end{minipage}}\end{center}
%
 %\caption{\it MapReduce algorithm for PageRank vertex ranking, using
 %4 MapReduce objects $A$, $MT$, $x$, and $y$.}
%
 %\label{fig:pr}
%\end{figure}

@@

$*$ Pre-aggregated MapReduce object $A$ = $|V| \times |V|$ matrix:  Key = $i$, Value = $[j, A_{ij}]$ \\
$*$ Pre-aggregated MapReduce object $MT$ = indices of all-zero rows of $A$:  Key = $i$, Value = 0 \\
$*$ Pre-aggregated MapReduce object $x$ = initial PageRank vector:  Key = $i$, Value = $1/|V|$ \\

While (${residual} > {tolerance}$) \\
\> 1 Compute contribution for random jumps and zero-outdegree vertices \\
\> \> Add: $x$ to $MT$ \\
$*$\> \> Convert $MT$:  Vertex is Key \\
\> \> Reduce $MT$:   \\
\> \> \> Input:  Key = $i$; MultiValue = $x_i, 0$ if $i \in {MT}$; MultiValue = $x_i$ if $i \notin {MT}$ \\
\> \> \> Compute:  Local contributions $c_{local}$ \\
\> \> \> Output:  If ${nvalues} = 2$, Key = $i$; Value = 0 \\
\> \> MPI\_Allreduce $c_{local}$ with MPI\_SUM to compute global contribution $c_{global}$ \\
\> 2 Compute $y = A^T x$  \\
\> \> Copy $x$ to $y$ \\
\> \> Add $A$ to $y$ \\
$*$\> \> Convert $y$:  Key is row index of $A$ \\
\> \> Reduce $y$: \\
\> \> \>  Input: Key = $i$; Multivalue = $y_i, [j, A_{ij}]$ for nonzeros $A_{ij}$ in $A$ \\
\> \> \>  Output:  Key = $j$; Value = $A_{ij} y_i$ \\
\> \> Collate $y$:  Key is column index of $A$ \\
\> \> Reduce $y$:  \\
\> \> \> Input:  Key = $j$; Multivalue = $A_{ij} y_i$ for all non-zero entries $A_{ij}$ of column $j$ \\
\> \> \> Output:  Key = $j$; Value = $\sum_i A_{ij} y_i$ \\
\> 3 Add contribution $c_{global}$ to $y$; compute local max norm $n_{local}$ of $y$ \\
\> \> Map $y$:  \\
\> \> \> Input:  Key = $i$, Value = $y_i$ \\
\> \> \> Compute: $y_i = y_i + c_{global}$; if $y_i > {n_{local}}$, $n_{local} = y_i$ \\
\> \> \> Output:  Key = $i$, Value = $y_i$ \\
\> \> MPI\_Allreduce $n_{local}$ with MPI\_MAX to compute global max norm $n_{global}$ \\
\> 4 Scale $y$ by $n_{global}$ \\
\> \> Map $y$: \\
\> \> \> Input:  Key = $i$, Value = $y_i$ \\
\> \> \> Output:  Key = $i$, Value = $y_i / {n_{global}}$ \\
\> 5 Compute residual \\
\> \> Add $y$ to $x$ \\
$*$\> \> Convert $x$:  Key is index $i$ \\
\> \> Reduce $x$:   \\
\> \> \> Input:  Key = $i$; Multivalue = $x_i, y_i$ \\
\> \> \> Compute:  if $|x_i - y_i| > {resid}$, ${resid} = |x_i - y_i|$ \\
\> \> MPI\_Allreduce with MPI\_MAX to compute $residual$ \\
\> 6 $x = y$; \\
Output: PageRank vector $x$
  \end{tabbing}
 \end{minipage}}\end{center}

 \caption{\it MapReduce algorithm for PageRank vertex ranking, using 4
 MapReduce objects $A$, $MT$, $x$, and $y$.  Pre-aggregating several
 of the MapReduce objects allows many communication intensive {\it
 collate()} operations to be replaced by local {\it convert()}
 operations (marked with asterisks).}

@@

A triangle in a graph is any triplet of vertices $(V_i,V_j,V_k)$ where
the edges $(V_i,V_j), (V_j,V_k), (V_i,V_k)$ exist.  Figure
\ref{fig:tri} outlines a MapReduce algorithm that enumerates all
triangles, assuming an input graph $G$ of undirected edges $(V_i,V_j)$
where $V_i < V_j$ for every edge, i.e. an upper-triangular R-MAT
matrix.  This exposition follows the triangle-finding algorithm
presented in~\cite{Cohen09}.

@@

MapReduce object $G$ = edges of graph: Key = $(V_i,V_j)$ with $V_i < V_j$, Value = NULL \\
